P3692R0: How to Avoid OOTA Without Really Trying

### 2025-05-13 paper/ootavacuous.tex
ootavacuous: Add 6.4.3 discussion section
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
ootavacuous: Set up to be P3692R0
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{document}
\title{DNNNNR0: How to Avoid OOTA Without Really Trying}

@@

This paper is a condensation of P3064R2 ("How to Avoid OOTA Without
Really Trying")~\cite{PaulEMcKenney2023P3046R2/OOTAwithoutTrying},
which was criticized as being overly long.
It also adds citations and clarifications.

@@

Semantic dependencies form only one type of link in an OOTA cycle.
The other type extends from a given store to a load that returns the
value stored.
It is tempting to argue that the store must precede
the load in global time and combine this with the intuitive notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that OOTA cycles cannot occur.
The idea has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf~\cite{PaulEMcKenney2014OOTA},
% This citation credits the formulation to Ali Sezgin, third author
% of N4323 ("Out-of-Thin-Air Execution is Vacuous").
where sdep is the set of semantic dependencies within each thread and
rf is the set of store-to-load ``reads-from'' links, whether internally
within a thread (rfi) or externally between threads (rfe).\footnote{
	Additional background on temporal distinctions between coe, fre,
	and rfe maybe found elsewhere~\cite[Appendix A]{PaulEMcKenney2023P3046R2/OOTAwithoutTrying}.}

@@

@@

Here we consider some general questions related to our definition of
semantic dependency.
Additional information is available in a companion technical report,
which is not cited in order to maintain anonymity.
% P3064R2 ("How to Avoid OOTA Without Really Trying")
% https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2024/p3064r2.pdf

>>>>

\begin{document}
\title{P3692R0: How to Avoid OOTA Without Really Trying}

@@

This paper is a condensation of P3064R2 (``How to Avoid OOTA Without
Really Trying'')~\cite{PaulEMcKenney2023P3046R2/OOTAwithoutTrying},
which was criticized as being overly long.
It also adds citations and clarifications.

@@

Semantic dependencies form only one type of link in an OOTA cycle.
The other type extends from a given store to a load that returns the
value stored.
It is tempting to argue that the store must precede
the load in global time and combine this with the intuitive notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that OOTA cycles cannot occur.
The idea has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf~\cite{PaulEMcKenney2014OOTA},
% This citation credits the formulation to Ali Sezgin, third author
% of N4323 (``Out-of-Thin-Air Execution is Vacuous'').
where sdep is the set of semantic dependencies within each thread and
rf is the set of store-to-load ``reads-from'' links, whether internally
within a thread (rfi) or externally between threads (rfe).\footnote{
	Additional background on temporal distinctions between coe, fre,
	and rfe maybe found elsewhere~\cite[Appendix A]{PaulEMcKenney2023P3046R2/OOTAwithoutTrying}.}

@@

\subsubsection{Discussion}
\label{sec:Discussion}

Allowing the compiler to transform the source code to assembly code
prior to analysis provides great simplifications.
This enables us to show the main result, namely that current compiler-based
C++ implementations do not produce OOTA cycles.
Note well that these compilers' current analysis suffices.

@@

Here we consider some general questions related to our definition of
semantic dependency.
Additional information is available in a companion technical report,
which is not cited in order to maintain anonymity.
% P3064R2 (``How to Avoid OOTA Without Really Trying'')
% https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2024/p3064r2.pdf
